### 2019-02-26 paper/main.tex
latex

\title{Reconstruction of strange attractors via inter-spike intervals}
\author{\`Alex Arcas Cuerda \& Ram\'on Marc Garcia Seuma}
\date{\today}
\documentclass[10pt]{article}

\usepackage[a4paper, total={7in, 10in}]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}

\begin{document}
\maketitle
\section{Introduction}

\section{Geometry from a Time Series}

\section{Reconstruction of dynamical systems from inter-spike intervals}

\section{Results}

\section{Conclusions}

\end{document}